---
title: "Adyen Loan Test"
author: "Brandon Wong"
date: "23 January 2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown file for Technical Test

This is the final code for my technical test.
There are 7 questions in total.

## Libraries needed + Setup

```{r}
options(max.print=999999) # Increase limit for max.print, 800k+ rows were omitted. Then read CSV "loan.csv" file (stored locally in this case)
options(scipen=999) # Remove scientific notation, where ### is the switch point
Original_Loan_File <- read.csv("C://Users//Brandon Wong//Documents//Adyen_Test//wetransfer-937f2c//data-analysis-test//loan.csv", header = TRUE, sep = ",")

library(DescTools)
library(ggplot2)
library(dplyr)
```

## Questions:

### Question 1: Provide a distribution of loans 

#### Using DescTools package, one can view the 1. Density graph, 2. Box and Whisker plot, 3. Empirical CDF plot

```{r}
Desc(Original_Loan_File$loan_amnt, plotit = TRUE, main = "Applied Loan Amount")
```

### Question 2: What is the lowest loan amount? 

```{r}
min(Original_Loan_File$loan_amnt, na.rm = TRUE) # Note: This can also be seen from Qn 1 result
```

### Question 3: What is the highest loan amount?

```{r}
max(Original_Loan_File$loan_amnt, na.rm = TRUE) # Note: This can also be seen from Qn 1 result
```

### Question 4: Provide a graph that visualizes the issued loans over time

#### This is a graph of Loan Amounts over time

```{r}
Original_Loan_File$issue_d <- as.Date(gsub("^", "01-", Original_Loan_File$issue_d), format="%d-%b-%Y") # Format date

OverTime_df1 <- Original_Loan_File %>% 
  select(issue_d, loan_amnt) %>% 
  group_by(issue_d) %>% 
  summarise(Sum_Of_Amount = sum(loan_amnt))

#Plotting chart
issuedovertime_chart <- ggplot(OverTime_df1, 
                  aes(x = issue_d, y = Sum_Of_Amount))
issuedovertime_chart + geom_line() + ggtitle("Issued Loans Over Time") + labs(x = "Loan Date Issued", y = "Loan Amount")
```

#### This is a more detailed graph of Loan Amounts over time (by Grade)

```{r}
OverTime_df2 <- Original_Loan_File %>% 
  select(issue_d, loan_amnt, grade) %>% 
  group_by(issue_d, grade) %>% 
  summarise(Sum_Of_Amount = sum(loan_amnt))

# Plotting chart
issuedovertimegrades_chart <- ggplot(OverTime_df2, aes(x = issue_d, y = Sum_Of_Amount))
issuedovertimegrades_chart + geom_area(aes(fill=grade)) + ggtitle("Issued Loans Over Time by Grade") + labs(x = "Date Issued", y = "Loan Amount") + scale_fill_discrete(name="Grades (A-G)")
```

### Question 5: Provide a graph that visualizes the loan status

```{r}
Desc(Original_Loan_File$loan_status, main = "Loan Status", plotit = T)

# This graph shows the distribution of loan amounts by status
LoanAmnt_by_Status_chart <- ggplot(Original_Loan_File, aes(loan_status, loan_amnt))
LoanAmnt_by_Status_chart + geom_boxplot(aes(fill = loan_status)) +
  theme(axis.text.x = element_blank()) +
  ggtitle("Loan Amount by Status") +
  labs(list(x = "Loan Status", y = "Loan Amount")) +
  scale_fill_discrete(name="Status Legend")
```

### Question 6: Provide a graph that visualizes the reason people took out a loan in percentages

```{r}
Desc(Original_Loan_File$purpose, main = "Loan Purposes", plotit = TRUE)
Desc(Original_Loan_File$title, main = "Loan Titles", plotit = TRUE)
```


### Question 7: Show other interesting insights 

#### The previous question caught my eye, where column "purpose" indicated why people took loans. As such, I sought to find out what the probability of unpaid loans was among the different loan purposes.

```{r}
Original_Loan_File$PaidVsUnpaid <- "Unpaid" # creating a new column and filling
Original_Loan_File$PaidVsUnpaid[which(Original_Loan_File$loan_status == "Fully Paid"
                                      |Original_Loan_File$loan_status == "Current"
                                      |Original_Loan_File$loan_status == "Does not meet the credit policy. Status:Fully Paid") ] <- "Paid or Current"

Original_Loan_File$PaidVsUnpaid <- factor(Original_Loan_File$PaidVsUnpaid)
Original_Loan_File$PaidVsUnpaid <- factor(Original_Loan_File$PaidVsUnpaid, levels = rev(levels(Original_Loan_File$PaidVsUnpaid)))
table(Original_Loan_File$PaidVsUnpaid) # Output of no. of Paid Loans vs Unpaid Loans
```

#### Now that the new variables are created, we can go ahead with the plotting of the chart

```{r}
Purpose_Of_Loan <- aggregate(loan_amnt ~ purpose + PaidVsUnpaid, 
                             data = Original_Loan_File, sum)

# This is the Probability of Unpaid Loans chart
ggplot(Purpose_Of_Loan, aes(x = purpose, y = loan_amnt, fill = PaidVsUnpaid)) + 
  ggtitle("Probability of Unpaid Loans") + 
  labs(x = "Purpose Of Loan", y = "Proportion of Loan Amount (%)") + 
  scale_fill_discrete(name="Paid vs Unpaid Loans") + 
  geom_bar(position = "fill", stat = "identity", colour = "black") + 
  theme(axis.text.x=element_text(size=15, angle = 90))
```

### This graph would go to show that educational loans are rather high risk, where almost a quarter (25%) of them are unpaid. Thus, the stakeholder might find it wise to limit the instances of educational loans.